\chapter{Introduction}
\label{ch:intro}
Nowadays Model Predictive Control (MPC) perhaps the most advanced control strategy available on the market. This method is based on optimization, thus mathematically provable, that  the calculated control inputs ensure the best control performance for the present scenario, which is a big advantage. Another, perhaps an even bigger advantage of MPC, is that different physical limitations of the process and constrains on measured variables, control variables, and controlled variables are a part of that optimization problem. And if the optimization problem is feasible, we can guarantee, that those constraints would not be violated during the plant operation. As it is a predictive control strategy, it uses accurate model predictions of the future behavior of the system and can provide early warnings of potential problems.

As we can see an MPC is a great control strategy. But nothing in this world is for free and MPC is not an exception. The MPC strategy is expensive, both in time and resources. MPC solves an optimization problem. This optimization problem could also be both equality and inequality constrained. And the presence of these constraints can make the optimization problem extremely difficult to solve. Therefore pretty strong computation machines are required. The MPC also requires an appropriate process model to compute accurate predictions. And the main industrial restriction on MPC is time. Most of the industrial problems are really complicated because of the high number of process and manipulated variables and their intercations. And the MPC must be able to solve such problems during only one sampling period.

This thesis is dedicated to designing a predictive controller, that deals with the main restrictions on MPC such as short available time to solve an optimization problem and nonlinear constraints. To test such MPC the Furuta Pendulum is used. Furuta pendulum is a rotational inverted pendulum and it is a great example of a fast, nonlinear process with complex dynamics. It was invented in 1992 at the Tokyo Institute of Technology by Katsuhisa Furuta as an example of a complex nonlinear oscillator, with the purpose of process control algorithm testing. This pendulum is underactuated and extremely non-linear due to gravitational forces and Coriolis interactions. As the main task of the thesis is to perform a swing-up control of the pendulum, two different approaches are used: first is a heuristic swing-up control, where the pendulum is controlled by two controllers: energy-shaping controller and a linear MPC. The second approach is an optimal swing-up control, where a Nonlinear Model Predictive Control (NMPC) strategy is used.

The thesis consists of two main parts: the first is a theoretical foundation, and the second illustrates simulation control of the pendulum device. In the first part operation of the pendulum is explained and its mathematical model derived. Also in that part, the theory for individual controllers is described. In the second part, both heuristic and optimal swing-up control approaches are explained as well as \textsc{MATLAB} implementation, and the results of the control are illustrated.   
